---
title: "NBA resale ticket prices"
runtime: shiny
output: 
  html_document: 
    theme: spacelab
---


I moved to the Bay Area almost three years ago, just in time to see the Golden State Warriors start to get pretty good. And then they got [really, really, good](https://www.youtube.com/watch?v=UdvA9w3Tmoc). Being the serious, hardworking PhD student that I am, following the Warriors became a hobby. Turns out that this hobby, [while healthier than drugs](http://www.cnn.com/2012/04/13/health/side-effects-sports-fan/), may not be any cheaper.

I grabbed all the data on resale tickets from [StubHub](https://www.stubhub.com/) for all remaining games of the 2015-16 season. As at 1 Feb, there were around 264,200 tickets for sale on StubHub. Unsuprisingly, going to Warriors home games is generally much more expensive than for any other team. At $285, the median ticket price for remaining games at Oracle Arena is around $100 more than the next closest team, the Spurs. 

The median home game ticket price generally descreases as you go down the current standings. The notable exception is the LA Lakers, where you can pay $280 to [probably see them lose](http://bleacherreport.com/articles/2613737-la-lakers-have-somehow-surpassed-philadelphia-76ers-as-nbas-worst). But then how long until we see another Kobe?


<center> <h4>Median ticket price for remaining 15/16 home games</h4> </center>

```{r, results = "asis", cache=F, comment=NA, echo=F}
library(rCharts)
library(knitr)

dsum.home <- read.csv("dsum.home.csv")
cols <- c("#006BB6", "#552582", "#000000", "#007DC3", "#F58426", "#860038", "#006BB6", 
          "#F9A01B", "#CE1141", "#C4CED3", "#00A94F", "#CE1141", "#4FA8FF", "#724C9F", 
          "#CE1142", "#E03A3E", "#F0163A", "#000000", "#23375B", "#008348", "#20385B", 
          "#1D1160", "#E56020", "#FFC633", "#002566", "#00471B", "#002B5C", "#B4975A", "#006BB6", "#001F70" )

standPlot <- nPlot(
  med.price ~ standing, 
  data = dsum.home, 
  group = "team",
  type = "scatterChart", showLegend=F)

# Add axis labels and format the tooltip
standPlot$yAxis(axisLabel = "Ticket price, home games ($)", width = 62, tickValues = "#! function (x) {    
                tickvalues = [25, 75, 125, 175, 225, 275, 325, 375];
                return tickvalues;
                } !#")

standPlot$xAxis(axisLabel = "Current Team Standing", 
                tickValues = "#! function (x) {    
                tickvalues = [0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30];
                return tickvalues;
                } !#")

standPlot$chart(color = cols, tooltipContent = "#! function(key, x, y){
        return '<h3>' + key + '</h3>' + 
        '<p>' + 'Standing: '+ x + ' , ' + 'Price: $'+ y + '</p>'
        } !#")


standPlot$chart(showLegend=F, sizeRange = c(250,250), forceY = c(25, 375), forceX = c(0, 30))

standPlot$setTemplate(
  afterScript = "<style>
.tick line {
  opacity: 0;
}
</style>")

standPlot$show('iframesrc',cdn=TRUE)
```


### Everyone wants to see Curry (+ the LeBron Effect)

Oracle is notorious for its [electric atmosphere](http://abc7news.com/sports/expert-decibel-level-at-oracle-is-similar-to-jet-engine/773850/), and Warriors fans are some of the most passionate in the league, so you may think there is a price premium for experiencing the home-court advantage. But while the home game prices for the Warriors [are enough the make you fall over](https://www.youtube.com/watch?v=LB4rplyj3TU), the away game prices are even worse. The median price is $340 for all remaining away games for the Warriors, $55 more than home games. Everyone wants to see [Curry score a three from their half court logo](https://www.youtube.com/watch?v=F8uoujg1XgE). 

For most other teams, home and away median prices are pretty close. The notable exception is the Cleveland Cavaliers, where the away game median price more than twice the home game price. This can only be explained at the LeBron Effect: the King's lands reach far and wide. 

<center> <h4>Median ticket price, home games versus away games</h4> </center>
```{r, results = "asis", cache=F, comment=NA, echo=F}
library(rCharts)
library(knitr)

dsum.home <- read.csv("dsum.home.csv")
cols <- c("#006BB6", "#552582", "#000000", "#007DC3", "#F58426", "#860038", "#006BB6", 
          "#F9A01B", "#CE1141", "#C4CED3", "#00A94F", "#CE1141", "#4FA8FF", "#724C9F", 
          "#CE1142", "#E03A3E", "#F0163A", "#000000", "#23375B", "#008348", "#20385B", 
          "#1D1160", "#E56020", "#FFC633", "#002566", "#00471B", "#002B5C", "#B4975A", "#006BB6", "#001F70" )

pricePlot <- nPlot(
med.price ~ away.med.price, 
  data = dsum.home, 
  group = "team",
  type = "scatterChart", showLegend=F)

# Add axis labels and format the tooltip
pricePlot$yAxis(axisLabel = "Ticket price, home games ($)", width = 62, tickValues = "#! function (x) {    
    tickvalues = [25, 75, 125, 175, 225, 275, 325, 375];
    return tickvalues;
} !#")

pricePlot$xAxis(axisLabel = "Ticket price, away games ($)", 
                tickValues = "#! function (x) {    
    tickvalues = [25, 75, 125, 175, 225, 275, 325, 375];
    return tickvalues;
} !#")

pricePlot$chart(color = cols, tooltipContent = "#! function(key, x, y){
        return '<h3>' + key + '</h3>' + 
        '<p>' + 'Away: $'+ x + ' , ' + 'Home: $'+ y + '</p>'
        } !#")

pricePlot$chart(showLegend=F, sizeRange = c(250,250), forceY = c(25, 375), forceX = c(25, 375))
pricePlot$setTemplate(
  afterScript = "<style>
.tick line {
  opacity: 0;
}
</style>")
#pricePlot$set(width = 600, height = 600)
pricePlot$show('iframesrc',cdn=TRUE)
```


In general, you pay for what you get: the higher the combined ranking of the two teams playing, the more expensive the tickets. Lakers games buck this trend. Note that for visibility I left the most expensive game off this chart: Kobe's final game at the Staples Center on 13 April has a median ticket price of [around $1650](https://www.stubhub.com/los-angeles-lakers-tickets-los-angeles-lakers-los-angeles-staples-center-4-13-2016/event/9370774/)!

Despite the priceness of Warrior away-games in general,  it could still be cheaper to [fly to Utah](https://www.kayak.com/flights/SFO-SLC/2016-03-30/2016-03-31) in late March to see the Jazz game, than to see any remaining home games. The most expensive games are between the Warriors and Spurs or OKC, the two teams who are most likely to threaten the Warriors chances of breaking the [Bull's 72-10 record](http://www.nba.com/history/season/19951996.html). Either way, Steve Kerr still may be the [luckiest man in NBA history](http://www.huffingtonpost.com/2015/06/17/steve-kerr-warriors-coach_n_7603384.html). 


<center> <h4>Median ticket price, all remaining NBA games</h4> </center>
```{r, echo=FALSE, cache = FALSE}
shinyAppDir("Shiny/app1", options=list(
    width="120%", height=500
  ))
```

#### Notes:
These charts were made using `rCharts` (thanks Ramnath Vaidyanathan for the awesome package). Code for scraping StubHub and analysis can be found here. 

